---
title: "Analysis Notebook"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r Load library}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(ggpubr))
```

```{r Load Data}
#Les_diffe_rents_accents_du_franc_ais_Asians_Whites_October_26_2022_08_24 <- read_excel("Data/Les différents accents du français - Asians+Whites_October 26, 2022_08.24.xlsx")
raw_data <- Les_diffe_rents_accents_du_franc_ais_Asians_Whites_October_26_2022_08_24


print(raw_data[1,])
```


```{r Rename Columns}
data <- raw_data[-1, ]

data <- data %>% 
  drop_na(`1A_1`)

data <- data %>%
  select(-c(IPAddress, Finished, ResponseId, QID18_1, QID18_2, QID18_3,
            QID18_5, QID18_6, QID18_7, QID18_8, QID18_9, QID18_10, QID19, `10B`, Courriel,
            `1B`, `2B`, `3B`, `4B`, `5B`, `6B`, `7B`, `8B`, `9B`, `10B`)) %>% 
  rename(gender = Q1,
         gender_other = Q1_4_TEXT, 
         age = Q2,
         country_of_birth = Q3,
         grow_up_province = Q41,
         grow_up_city = Q4,
         field_of_study = Q5,
         ethnic_group = Q6,
         ethnic_group_other = Q6_13_TEXT, 
         num_of_language = Q7...123, 
         spoken_language = Q7...124, 
         spoken_language_other = Q7_11_TEXT,
         mother_tongue = Q8,
         mother_tongue_other = Q8_11_TEXT,
         year_of_french = Q9,
         french_speaking_place = Q10,
         french_speaking_place_other = Q10_6_TEXT,
         month_live_in_fr_env = Q10B,
         pref_french = Q11,
         pref_french_other = Q11_3_TEXT,
         teacher_origin = Q12, 
         teacher_origin_other = Q12_3_TEXT
         ) %>% 
  mutate(age = as.numeric(age),
         num_of_language = as.numeric(num_of_language),
         year_of_french = as.numeric(year_of_french),
         )

# convert rating 
data <- data %>%  
  mutate_at(vars(1:90), as.numeric)

data <- data %>% 
  mutate_if(is.character, as.factor)
```


# Participant Portrait

```{r # of participants}
nrow(data)
```


```{r Participant gender}
data %>% 
  group_by(gender) %>% 
  tally() %>%  
  mutate(prop = scales::percent(n/nrow(data)))


data %>%  
  select(gender, gender_other) %>% 
  filter(gender %in% c('Prefer not to answer', 'Prefer to define myself', NA))
# 1 person recognize as agnder, others did not reveal 
```

```{r Participant Age}
range(data$age, na.rm = TRUE)

# 9 missing value  
data %>%  
  ggplot(aes(x = age)) + 
  geom_density() + 
  geom_vline(xintercept = mean(data$age, na.rm = TRUE), color = 'red') + 
  theme_bw()
```

```{r Country of Birth}
# born in Canda
data %>%  
  select(country_of_birth) %>%  
  mutate(Canada = ifelse(country_of_birth == 'Canada', TRUE, FALSE)) %>% 
  group_by(Canada) %>% 
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))


data %>%  
  drop_na(country_of_birth) %>% 
  select(country_of_birth) %>% 
  filter(country_of_birth != 'Canada') %>% 
  mutate(continent = case_when(country_of_birth %in% c('China', 'India', 'South Korea', 'Japan', 'Pakistan', 'Philippines') ~ 'Asia', 
                               country_of_birth %in% c('Mauritius', 'Egypt', 'Nigeria') ~ 'Africa',
                               country_of_birth %in% c('United Kingdom of Great Britain and Northern Ireland', 'Czech Republic', 'Italy', 'Germany') ~ 'Europe',
                               country_of_birth %in% c('United States of America', 'Brazil', 'Mexico') ~ 'America')) %>%  
  group_by(continent) %>%  
  tally()




```

```{r Grow-up Province}
data %>% 
  filter(country_of_birth == 'Canada') %>%  
  group_by(grow_up_province) %>% 
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))
```

```{r Field of Study}
data %>%  
  group_by(field_of_study) %>% 
  tally() %>% 
  filter(field_of_study %in% c('Français', 'French', 'Linguistic', 'linguistique et psychologie', 'linguistiques', 'Linguistiques (Arts)'))
# six people are with French or linguitistic background
```
```{r ethnic group}
data %>% 
  group_by(ethnic_group) %>% 
  tally()
```




```{r # of Language Spoken}
# 13 missing data 
data %>% 
  # drop_na(num_of_language) %>% 
  group_by(num_of_language) %>% 
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n))) 
```


```{r Spoken Language}
# 7 missing values 
data %>% 
  drop_na(spoken_language) %>% 
  select(spoken_language, spoken_language_other) %>% 
  mutate(English = ifelse(str_detect(spoken_language, 'English'), TRUE, FALSE)) %>% 
  filter(English != TRUE)
# 2 people don't speak English 
```

```{r Month Tongue}
month_tongue_df <- data %>% 
  select(mother_tongue) %>%
  mutate(English = ifelse(str_detect(mother_tongue, "English") | str_detect(mother_tongue, "Anglais"), TRUE, FALSE),
         Cantonese = ifelse(str_detect(mother_tongue,"Cantonese"), TRUE, FALSE), 
         Hindi = ifelse(str_detect(mother_tongue,"Hindi"), TRUE, FALSE),
         Mandarin = ifelse(str_detect(mother_tongue,"Mandarin"), TRUE, FALSE),      
         Korean = ifelse(str_detect(mother_tongue,"Korean"), TRUE, FALSE),
         Spanish = ifelse(str_detect(mother_tongue,"Spanish"), TRUE, FALSE),  
         German = ifelse(str_detect(mother_tongue,"German"), TRUE, FALSE),  
         Tagalog = ifelse(str_detect(mother_tongue,"Tagalog"), TRUE, FALSE))  



language_count = matrix(NA, 7, 2)
language_count[,1] <- names(month_tongue_df)[2:8]
for (i in 1:7){
  language_count[i,2] = as.numeric(table(month_tongue_df[,i+1]))[2]
}


language_count <- as.data.frame(language_count)
colnames(language_count) = c('mother_tongue', 'n')

language_count %>% 
  arrange(desc(n)) %>% 
  mutate(prop = scales::percent(n/sum(n)))
```

```{r Year of Studying French}

# 15 missing value
data %>% 
  # drop_na(year_of_french) %>% 
  mutate(year_interval = case_when(year_of_french<=5~'0-5', 
                                   year_of_french>5 & year_of_french <= 10~ '5-10', 
                                   year_of_french>10 & year_of_french<= 15~ '10-15', 
                                   year_of_french > 15~'15+')) %>% 
  group_by(year_interval) %>%
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))

  
  
  
  # drop_
  # tally() %>% 
  # ggplot(aes(x="", y = n, fill=year_group))+
  # geom_bar(width = 1, stat = "identity")+
  # coord_polar("y") + 
  # scale_fill_discrete(name = "Years of Studying French") 
```



```{r Live in French-speaking Area }
live_in_french_env_df <- as.data.frame(t(as.data.frame(data %>% 
  select(french_speaking_place)  %>% 
  group_by(french_speaking_place) %>% 
  tally() %>% 
  mutate(France = ifelse(french_speaking_place == 'France' | french_speaking_place == 'France, Elsewhere? if so please specify:', n, NA),
         Quebec = ifelse(french_speaking_place == 'Quebec' | french_speaking_place == 'Quebec, Elsewhere? if so please specify:', n, NA),
         never = ifelse(str_detect(french_speaking_place, 'I have never stayed in a French-speaking context.'), n, NA),
         other = ifelse(french_speaking_place == 'Elsewhere? if so please specify:', n, NA),
         both = ifelse(french_speaking_place %in% c('Quebec and France', 'Quebec, France'), n, NA),
         unkown = ifelse(is.na(french_speaking_place), n, NA)
         )%>% 
  summarize(France = sum(France, na.rm = TRUE),
            Quebec = sum(Quebec, na.rm = TRUE),
            never = sum(never, na.rm = TRUE),
            other = sum(other, na.rm = TRUE),
            both = sum(both, na.rm = TRUE), 
            unkown = sum(unkown, na.rm = TRUE)))))

live_in_french_env_df <- tibble::rownames_to_column(live_in_french_env_df, "Area")
live_in_french_env_df %>% 
  mutate(prop = scales::percent(V1/sum(V1),1))
```
```{r Months of Living in French-speaking areas}
levels(data$month_live_in_fr_env) <- c("1 - 3", "4 - 6", "7 - 12", "< 1", "> 12")
data$month_live_in_fr_env <- factor(data$month_live_in_fr_env, levels = c("< 1", "1 - 3", "4 - 6", "7 - 12", "> 12")) 
data %>% 
  group_by(month_live_in_fr_env) %>% 
  tally() %>%  
  mutate(prop = scales::percent(n/sum(n))) 
```

```{r Preferred Type of French}
data %>% 
  select(pref_french) %>% 
  group_by(pref_french) %>%
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))



```


```{r Origin of Teacher}
data %>% 
  group_by(teacher_origin) %>%  
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))

data %>%  
  filter(teacher_origin == 'Other(s), specify:') %>% 
  select(teacher_origin_other)
```


# Part I - Recording Evaluation: Accent

1A: Quebec. White.
2A: French. Asian.
3A. Acadian. Asian.
4A. L2 French. White.
5A. African. Asian. 
6A. Quebec. Asian.
7A. African. White. 
8A. Acadian. White. 
9A. French. White. 
10A. L2 French. Asian.




* General Perspective  
1) I think this is beautiful French 


* Understandability
2) Is it rather easy or difficult to understand the speaker?
3) Would the person who did the recording be a good French teacher at UBC?

* Solidarity dimension
5) This person is dynamic.
6) This person is nice.
7) This person is social.

* Status traits
8) This person is professional.
9) This person is leader.
10) This person is educated.


```{r Keep Recording Evluation Columns}
recording_df <- data %>% 
  select(c(1:90)) 


```



## General Perspective
```{r Beautiful French}
beautiful_fr <- 
  recording_df %>% 
  select(seq(1, 82, 9)) %>% 
  transmute(Quebec = `1A_1` + `6A_1`, 
         European = `2A_1` + `9A_1`, 
         Acadian = `3A_1` + `8A_1`, 
         `L2 French` = `4A_1` + `10A_1`, 
         African = `5A_1` + `7A_1`
        ) %>% 
  mutate(id = 1:nrow(recording_df))


# short to long 
beautiful_fr_long <- beautiful_fr %>%  
  gather(key = "Accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, Accent)


beautiful_fr_long %>%  
  group_by(Accent) %>% 
  get_summary_stats(value, type = 'mean_sd')
# Note: only 83 people complete the survey



# repeated measure ANOVA
anova_beau_fr <-
  get_anova_table(anova_test(
    data = beautiful_fr_long,
    dv = value,
    wid = id,
    within = Accent
  ))

# paired t test
beautiful_fr_pw <- beautiful_fr_long %>%
  pairwise_t_test(value ~ Accent,
                  paired = TRUE,
                  p.adjust.method = 'bonferroni') %>% 
  mutate(y.position = c(seq(12, 16.5, by = 0.5)))


p_beautiful <- ggboxplot(beautiful_fr_long, x = "Accent", y = "value", fill = "Accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(beautiful_fr_pw, vjust = 0.1, bracket.nudge.y = 0.6, size = 2,hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(anova_beau_fr, detailed = TRUE),
    caption = get_pwc_label(beautiful_fr_pw),
    title = 'Beautiful French',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))

```



```{r General Perspetive - race}
beatiful_fr_race <- 
  recording_df %>% 
  select(seq(1, 82, 9)) %>% 
  transmute(White = `1A_1` + `4A_1` +`7A_1` +`8A_1` +`9A_1`, 
            Asian = `2A_1` + `3A_1` +`5A_1` +`6A_1` +`10A_1`) %>% 
  mutate(id = 1:nrow(recording_df)) 


beatiful_fr_race %>% 
  summarise(across(White:Asian, mean, na.rm = TRUE))


# one-sided t-test
t.test(beatiful_fr_race$White, beatiful_fr_race$Asian, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(beatiful_fr_race$White, beatiful_fr_race$Asian, alternative= "two.side", paired = TRUE)
```



## Understandability 
```{r Understandability - Create df}
# 2) This person is easy to understand.
# 3) This person could be a good French teacher at UBC.

und_df <- recording_df %>%
  select(contains(c('_2', '3'))) %>%
      transmute(
        que_easy_und = `1A_2` + `6A_2`,
        fr_easy_und = `2A_2` + `9A_2`,
        aca_easy_und = `3A_2` + `8A_2`,
        l2_easy_und = `4A_2` + `10A_2`,
        afr_easy_und = `5A_2` + `7A_2`,
        que_good_teacher = `1A_3` + `6A_3`,
        fr_good_teacher = `2A_3` + `9A_3`,
        aca_good_teacher = `3A_3` + `8A_3`,
        l2_good_teacher = `4A_3` + `10A_3`,
        afr_good_teacher = `5A_3` + `7A_3`)
           
easy_und_df_long <- und_df %>% 
  select(que_easy_und, fr_easy_und, aca_easy_und, l2_easy_und, afr_easy_und) %>%
  rename(Quebec = que_easy_und,
         European = fr_easy_und,
         Acadian = aca_easy_und,
         `L2 French` = l2_easy_und,
         African = afr_easy_und
         ) %>% 
  mutate(id = 1: nrow(und_df)) %>% 
    gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)


good_teacher_df_long <- und_df %>% 
  select(que_good_teacher, fr_good_teacher, aca_good_teacher,l2_good_teacher, afr_good_teacher) %>%
  rename(Quebec = que_good_teacher,
         European = fr_good_teacher,
         Acadian = aca_good_teacher,
         `L2 French` = l2_good_teacher,
         African = afr_good_teacher) %>% 
  mutate(id = 1: nrow(und_df)) %>% 
    gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)  
```

```{r Understandability - race}
und_df_race <- recording_df %>%
  select(contains(c('_2', '3'))) %>%
      transmute(
        white_easy_und = `1A_2` + `4A_2` + `7A_2` + `8A_2` + `9A_2`,
        asian_easy_und = `2A_2` + `3A_2` + `5A_2` + `6A_2` + `10A_2`,
        white_good_teacher = `1A_3` + `4A_3` + `7A_3` + `8A_3` + `9A_3`,
        asian_good_teacher = `2A_3` + `3A_3` + `5A_3` + `6A_3` + `10A_3`) %>% 
  mutate(id = 1:nrow(recording_df)) 


und_df_race %>% 
  summarise(across(white_easy_und:asian_good_teacher, mean, na.rm = TRUE))


# one-sided t-test
t.test(und_df_race$white_easy_und, und_df_race$asian_easy_und, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(und_df_race$white_easy_und, und_df_race$asian_easy_und, alternative= "two.side", paired = TRUE)
```


```{r Understandability - Easy to understand}
easy_und_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_easy_und <- get_anova_table(anova_test(data = easy_und_df_long, dv = value, wid = id, within = accent))
  
easy_und_pw <- easy_und_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 

p_easy_und <- ggboxplot(easy_und_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(easy_und_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_easy_und, detailed = TRUE),
    caption = get_pwc_label(easy_und_pw),
    y = 'Score',
    title = 'Easy to Understand'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```
```{r Understandability - Good teacher - race}
  # short to long 
good_teacher_long_race <- und_df_race %>%  
  select(white_good_teacher, asian_good_teacher, id) %>% 
  rename(White = white_good_teacher,
         Asian = asian_good_teacher) %>% 
  gather(key = "Race", value = "value", White, Asian) %>%
  convert_as_factor(id, Race)


good_teacher_long_race %>%  
  group_by(Race) %>% 
  get_summary_stats(value, type = 'mean_sd')
# Note: only 83 people complete the survey


# one-sided t-test
t.test(und_df_race$white_good_teacher, und_df_race$asian_good_teacher, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(und_df_race$white_good_teacher, und_df_race$asian_good_teacher, alternative= "two.side", paired = TRUE)
```


```{r Understandability - Good teacher}
good_teacher_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_good_teacher <- get_anova_table(anova_test(data = good_teacher_df_long, dv = value, wid = id, within = accent))
  
good_teacher_pw <- good_teacher_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 

p_good_teacher <- ggboxplot(good_teacher_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(good_teacher_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_good_teacher, detailed = TRUE),
    caption = get_pwc_label(good_teacher_pw),
    y = 'Score',
    title = 'Good French Teacher'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```


```{r Understandability - Plot, fig.height= 4, fig.width= 8}
cowplot::plot_grid(p_easy_und, p_good_teacher)

```
```{r Understandability - table}

matrix <- function(x, a,b) {
   base::matrix(x,nrow =a, ncol=b)
}

res_tb_und <- matrix(NA,  2, 7)

res_tb_und[1, 1:5] <- table_filler(easy_und_df_long)
res_tb_und[2, 1:5] <- table_filler(good_teacher_df_long)

res_tb_und[, 6] <- c(res_easy_und$F, res_good_teacher$F)
res_tb_und[, 7] <- c(res_easy_und$p, res_good_teacher$p)


res_tb_und <- as.data.frame(res_tb_und)
colnames(res_tb_und) <- c('mean_Acadian', 'mean_African', 'mean_European','mean_L2', 'mean_Quebec', 'F', 'p-value')
rownames(res_tb_und) <- c('Easy to Understand', 'Good French Teacher')
res_tb_und

```


## Solidarity dimension

```{r Solidarity dimension - create df}
# 5) This person is dynamic.
# 6) This person is nice.
# 7) This person is social.

sd_df <- recording_df %>% 
  select(contains(c('_5', '_6', '_7'))) %>%
  transmute(que_dynamic = `1A_5` + `6A_5`, 
         fr_dynamic = `2A_5` + `9A_5`, 
         aca_dynamic = `3A_5` + `8A_5`, 
         l2_dynamic = `4A_5` + `10A_5`, 
         afr_dynamic = `5A_5` + `7A_5`, 
         que_nice = `1A_6` + `6A_6`, 
         fr_nice = `2A_6` + `9A_6`, 
         aca_nice = `3A_6` + `8A_6`, 
         l2_nice = `4A_6` + `10A_6`, 
         afr_nice = `5A_6` + `7A_6`, 
         que_social = `1A_7` + `6A_7`, 
         fr_social = `2A_7` + `9A_7`, 
         aca_social = `3A_7` + `8A_7`, 
         l2_social = `4A_7` + `10A_7`, 
         afr_social = `5A_7` + `7A_7`, 
          )

# dynamic 
dynamic_df_long <- sd_df %>% 
  select(c(1:5)) %>%  
  rename(Quebec = que_dynamic,
         European = fr_dynamic,
         Acadian = aca_dynamic,
         `L2 French` = l2_dynamic,
         African = afr_dynamic
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)


# nice 
nice_df_long <- sd_df %>% 
  select(c(6:10)) %>%  
  rename(Quebec = que_nice,
         European = fr_nice,
         Acadian = aca_nice, 
         `L2 French` = l2_nice,
         African = afr_nice
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)

# social 

social_df_long <- sd_df %>% 
  select(c(11:15)) %>%  
  rename(Quebec = que_social,
         European = fr_social,
         Acadian = aca_social,
         `L2 French` = l2_social,
         African = afr_social
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)
```

```{r Solidarity dimension - race}
sd_race_df <-recording_df %>% 
  select(contains(c('_5', '_6', '_7'))) %>%
  transmute(
    white_dynamic = `1A_5` + `4A_5` + `7A_5` + `8A_5`+ `9A_5`, 
    asian_dynamic = `2A_5` + `3A_5` + `5A_5` + `6A_5`+ `10A_5`, 
    white_nice = `1A_6` + `4A_6` + `7A_6` + `8A_6`+ `9A_6`, 
    asian_nice = `2A_6` + `3A_6` + `5A_6` + `6A_6`+ `10A_6`, 
    white_social = `1A_7` + `4A_7` + `7A_7` + `8A_7`+ `9A_7`, 
    asian_social = `2A_7` + `3A_7` + `5A_7` + `6A_7`+ `10A_7`
          )


sd_race_df %>% 
  summarise(across(white_dynamic:asian_social, mean, na.rm = TRUE))

## Dynamic 
# one-sided t-test
t.test(sd_race_df$white_dynamic, sd_race_df$asian_dynamic, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(sd_race_df$white_dynamic, sd_race_df$asian_dynamic, alternative= "two.side", paired = TRUE)


# Nice
# one-sided t-test
t.test(sd_race_df$white_nice, sd_race_df$asian_nice, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(sd_race_df$white_nice, sd_race_df$asian_nice, alternative= "two.side", paired = TRUE)


# Social 
# one-sided t-test
t.test(sd_race_df$white_social, sd_race_df$asian_social, alternative= "two.side", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(sd_race_df$white_social, sd_race_df$asian_social, alternative= "two.side", paired = TRUE)
```






```{r SD - Dynamic}
dynamic_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_dynamic <- get_anova_table(anova_test(data = dynamic_df_long, dv = value, wid = id, within = accent))
  
dynamic_pw <- dynamic_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 


p_dynamic <- ggboxplot(dynamic_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(dynamic_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_dynamic, detailed = TRUE),
    caption = get_pwc_label(dynamic_pw),
    y = 'Score',
    title = 'Dynamic'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```
```{r SD - Nice}
nice_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  

res_nice <- get_anova_table(anova_test(data = nice_df_long, dv = value, wid = id, within = accent))
  
nice_pw <- nice_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1)))
p_nice <- ggboxplot(nice_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(nice_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_nice, detailed = TRUE),
    caption = get_pwc_label(nice_pw),
    title = 'Nice',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))

```

```{r SD - Social}
social_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_social <- get_anova_table(anova_test(data = social_df_long, dv = value, wid = id, within = accent))
  
social_pw <- social_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1)))
(p_social <- ggboxplot(social_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(social_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res_social, detailed = TRUE),
    caption = get_pwc_label(social_pw),
    title = 'Social',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold')))
```

```{r SD Plot, fig.height= 8, fig.width= 8}
cowplot::plot_grid(p_beautiful,p_dynamic, p_nice, p_social)
```

```{r SD Table}
table_filler <- function(df) {
  res <- pull(df %>%
    group_by(accent) %>%
    summarize(mean = round(mean(value, na.rm = TRUE), 1)), mean)
  return(res)
}

# matrix <- function(x, a,b) {
#    base::matrix(x,nrow =a, ncol=b)
# }

res_tb <- matrix(NA ,3, 7)

res_tb[1, 1:5]  <- table_filler(dynamic_df_long)
res_tb[2, 1:5]  <- table_filler(nice_df_long)
res_tb[3, 1:5]  <- table_filler(social_df_long)


res_tb[, 6] <- c(res_dynamic$F, res_nice$F, res_social$F)
res_tb[, 7] <- c(res_dynamic$p, res_nice$p, res_social$p)

res_tb <- as.data.frame(res_tb)
colnames(res_tb) <-
  c(
    'mean_Acadian',
    'mean_African',
    'mean_L2',
    'mean_French',
    'mean_Quebec',
    'F',
    'p-value'
  )
rownames(res_tb) <- c('Dynamic', 'Nice', 'Social')
res_tb
```


## Status traits



```{r ST: create df}

# 8) This person is professional
# 9) This person is a good leader. 
# 10) This person is educated.


st_df <- recording_df %>%
  select(contains(c('_8', '_9', '_10'))) %>%
  transmute(
    que_prof = `1A_8` + `6A_8`,
    fr_prof = `2A_8` + `9A_8`,
    aca_prof = `3A_8` + `8A_8`,
    l2_prof = `4A_8` + `10A_8`,
    afr_prof = `5A_8` + `7A_8`,
    que_lead = `1A_9` + `6A_9`,
    fr_lead = `2A_9` + `9A_9`,
    aca_lead = `3A_9` + `8A_9`,
    l2_lead = `4A_9` + `10A_9`,
    afr_lead = `5A_9` + `7A_9`,
    que_edu = `1A_10` + `6A_10`,
    fr_edu = `2A_10` + `9A_10`,
    aca_edu = `3A_10` + `8A_10`,
    l2_edu = `4A_10` + `10A_10`,
    afr_edu = `5A_10` + `7A_10`,
  )


# professional  
prof_df_long <- st_df %>% 
  select(c(1:5)) %>%  
  rename(Quebec = que_prof,
         European = fr_prof,
         Acadian = aca_prof,
         `L2 French` = l2_prof,
         African = afr_prof
         ) %>% 
  mutate(id = 1: nrow(st_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)

# leadership  
lead_df_long <- st_df %>% 
  select(c(6:10)) %>%  
  rename(Quebec = que_lead,
         European = fr_lead,
         Acadian = aca_lead,
         `L2 French` = l2_lead,
         African = afr_lead
         ) %>% 
  mutate(id = 1: nrow(st_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)


# educated   
edu_df_long <- st_df %>% 
  select(c(11:15)) %>%  
  rename(Quebec = que_edu,
         European = fr_edu,
         Acadian = aca_edu,
         `L2 French` = l2_edu,
         African = afr_edu
         ) %>% 
  mutate(id = 1: nrow(st_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)



```


```{r}
st_race_df <-recording_df %>% 
  select(contains(c('_8', '_9', '_10'))) %>%
  transmute(
    white_prof = `1A_8` + `4A_8` + `7A_8` + `8A_8`+ `9A_8`, 
    asian_prof= `2A_8` + `3A_8` + `5A_8` + `6A_8`+ `10A_8`, 
    white_leader = `1A_9` + `4A_9` + `7A_9` + `8A_9`+ `9A_9`, 
    asian_leader = `2A_9` + `3A_9` + `5A_9` + `6A_9`+ `10A_9`, 
    white_edu = `1A_10` + `4A_10` + `7A_10` + `8A_10`+ `9A_10`, 
    asian_edu = `2A_10` + `3A_10` + `5A_10` + `6A_10`+ `10A_10`)
          


st_race_df %>% 
  summarise(across(white_prof:asian_edu, mean, na.rm = TRUE))

# two-sided t-test
t.test(st_race_df$white_prof, st_race_df$asian_prof, alternative= "two.side", paired = TRUE) 

# two-sided Wilcoxon test
wilcox.test(st_race_df$white_prof, st_race_df$asian_prof, alternative= "two.side", paired = TRUE)


# two-sided t-test
t.test(st_race_df$white_leader, st_race_df$asian_leader, alternative= "two.side", paired = TRUE) 

# two-sided Wilcoxon test
wilcox.test(st_race_df$white_leader, st_race_df$asian_leader, alternative= "two.side", paired = TRUE)


# one-sided t-test
t.test(st_race_df$white_leader, st_race_df$asian_leader, alternative= "less", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(st_race_df$white_leader, st_race_df$asian_leader, alternative= "less", paired = TRUE)


# two-sided t-test
t.test(st_race_df$white_edu, st_race_df$asian_edu, alternative= "two.side", paired = TRUE) 

# two-sided Wilcoxon test
wilcox.test(st_race_df$white_edu, st_race_df$asian_edu, alternative= "two.side", paired = TRUE) 
```



```{r ST - Professional}

prof_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_prof <- get_anova_table(anova_test(data = prof_df_long, dv = value, wid = id, within = accent))
  
prof_pw <- prof_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 
p_prof <- ggboxplot(prof_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(prof_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_prof, detailed = TRUE),
    caption = get_pwc_label(prof_pw),
    y = 'Score',
    title = 'Professional'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))

```

```{r ST - Leadership}
lead_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_lead <- get_anova_table(anova_test(data = lead_df_long, dv = value, wid = id, within = accent))
  
prof_pw <- lead_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 
p_lead <- ggboxplot(lead_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(prof_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_lead, detailed = TRUE),
    caption = get_pwc_label(prof_pw),
    y = 'Score',
    title = 'Good leader'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```
```{r ST - Educated}
edu_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_edu <- get_anova_table(anova_test(data = edu_df_long, dv = value, wid = id, within = accent))
  
prof_pw <- edu_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 
p_edu <- ggboxplot(edu_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(prof_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_edu, detailed = TRUE),
    caption = get_pwc_label(prof_pw),
    y = 'Score',
    title = 'Educated'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```

```{r ST - Plot, fig.height= 8, fig.width= 8}
cowplot::plot_grid(p_prof,p_lead, p_edu)

```
```{r ST - Table}
res_tb_st <- matrix(NA, 3, 7)

res_tb_st[1, 1:5] <- table_filler(prof_df_long)
res_tb_st[2, 1:5] <- table_filler(lead_df_long)
res_tb_st[3, 1:5] <- table_filler(edu_df_long)


res_tb_st[, 6] <- c(res_prof$F, res_lead$F, res_edu$F)
res_tb_st[, 7] <- c(res_prof$p, res_lead$p, res_edu$p)
res_tb_st <- as.data.frame(res_tb_st)
colnames(res_tb_st) <- c('mean_Acadian', 'mean_African', 'mean_l2','mean_French', 'mean_Quebec', 'F', 'p-value')
rownames(res_tb_st) <- c('Professional', 'Good Leader', 'Educated')
res_tb_st
```




# Part II: Recording Evaluation: Speaker Race and Accent



1A: Quebec. White.
2A: French. Asian.
3A. Acadian. Asian.
4A. L2 French. White.
5A. African. Asian. 
6A. Quebec. Asian.
7A. African. White. 
8A. Acadian. White. 
9A. French. White. 
10A. L2 French. Asian.


```{r Create df}
recording_df_agg <- recording_df %>%  
  mutate(id = 1:nrow(recording_df)) %>% 
  transmute(
    rec_1 = rowSums(.[2:10]),
    rec_2 = rowSums(.[11:19]),
    rec_3 = rowSums(.[20:28]),
    rec_4 = rowSums(.[29:37]),
    rec_5 = rowSums(.[38:46]),
    rec_6 = rowSums(.[47:55]),
    rec_7 = rowSums(.[56:64]),
    rec_8 = rowSums(.[65:73]),
    rec_9 = rowSums(.[74:82]),
    rec_10 = rowSums(.[83:91])
  ) %>%  
  mutate(id = 1:nrow(recording_df)) %>% 
  relocate(id, .before = rec_1) 


recording_df_agg_long <- recording_df_agg %>%  
  mutate(id = as.factor(id)) %>% 
  pivot_longer(-c(id), values_to = "score", names_to = "rec_index") %>% 
  mutate(speaker_race = ifelse(rec_index %in% c('rec_1', 'rec_4', 'rec_7', 'rec_8', 'rec_9'), 'white', 'asian'), 
         speaker_accent = case_when(rec_index %in% c('rec_1', 'rec_6') ~ 'Quebec', 
                                    rec_index %in% c('rec_2', 'rec_9') ~ 'European',
                                    rec_index %in% c('rec_3', 'rec_8') ~ 'Acadian',
                                    rec_index %in% c('rec_4', 'rec_10') ~ 'L2',
                                    rec_index %in% c('rec_5', 'rec_7') ~ 'African'))
  
```



### two-smaple test on speaker race

```{r two sample test}
w_df <- recording_df_agg_long %>%
  filter(speaker_race == "white")
# white speakers
a_df <- recording_df_agg_long %>%
  filter(speaker_race == "asian")

# one-sided t-test
t.test(w_df$score, a_df$score, alternative= "less", paired = TRUE) 

# one-sided Wilcoxon test
wilcox.test(w_df$score, a_df$score, alternative= "less", paired = TRUE)
```
```{r Two-way repeated measures ANOVA, warning=FALSE}
res.aov.twoway <- anova_test(
  data = recording_df_agg_long, dv = score, wid = id,
  within = c(speaker_accent, speaker_race)
  )
get_anova_table(res.aov.twoway)  
```



```{r Two-way repeated measures ANOVA - result}

# pairwise t-test
pwc <- recording_df_agg_long %>%
  group_by(speaker_accent) %>%
  pairwise_t_test(
    score ~ speaker_race, paired = TRUE,
    p.adjust.method = "bonferroni",
    alternative = 'greater'
    )


pwc <- pwc %>% add_xy_position(x = "speaker_accent")
ggboxplot(
  recording_df_agg_long, x = "speaker_accent", y = "score",
  color = "speaker_race", palette = "jco"
  ) + 
  stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.aov.twoway, detailed = TRUE),
    caption = get_pwc_label(pwc),
    y = 'Score',
    x = 'Speaker Accent'
  ) + 
  theme(axis.title.x = element_text(size = 15,face = 'bold'),
        axis.title.y = element_text(size = 15,face = 'bold'),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15))
pwc
```

```{r Examine the group of European and English}
recording_df_long %>% 
  filter(accent == 'European' |
         accent == 'English') %>% 
  group_by(accent, race) %>%
  summarise(mu = mean(score),
            sd = sd(score))
```




```{r }
names(recording_df)
```

#
















