---
title: "Analysis Notebook"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r Load library}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(ggpubr))
```

```{r Load Data}
#Les_diffe_rents_accents_du_franc_ais_Asians_Whites_October_26_2022_08_24 <- read_excel("Data/Les différents accents du français - Asians+Whites_October 26, 2022_08.24.xlsx")
raw_data <- Les_diffe_rents_accents_du_franc_ais_Asians_Whites_October_26_2022_08_24


print(raw_data[1,])
```


```{r Rename Columns}
data <- raw_data[-1, ]

data <- data %>%
  select(-c(IPAddress, Finished, ResponseId, QID18_1, QID18_2, QID18_3,
            QID18_5, QID18_6, QID18_7, QID18_8, QID18_9, QID18_10, QID19, `10B`, Courriel,
            `1B`, `2B`, `3B`, `4B`, `5B`, `6B`, `7B`, `8B`, `9B`, `10B`)) %>% 
  rename(gender = Q1,
         gender_other = Q1_4_TEXT, 
         age = Q2,
         country_of_birth = Q3,
         grow_up_province = Q41,
         grow_up_city = Q4,
         field_of_study = Q5,
         ethnic_group = Q6,
         ethnic_group_other = Q6_13_TEXT, 
         num_of_language = Q7...123, 
         spoken_language = Q7...124, 
         spoken_language_other = Q7_11_TEXT,
         mother_tongue = Q8,
         mother_tongue_other = Q8_11_TEXT,
         year_of_french = Q9,
         french_speaking_place = Q10,
         french_speaking_place_other = Q10_6_TEXT,
         month_live_in_fr_env = Q10B,
         pref_french = Q11,
         pref_french_other = Q11_3_TEXT,
         teacher_origin = Q12, 
         teacher_origin_other = Q12_3_TEXT
         ) %>% 
  mutate(age = as.numeric(age),
         num_of_language = as.numeric(num_of_language),
         year_of_french = as.numeric(year_of_french),
         )

# convert rating 
data <- data %>%  
  mutate_at(vars(1:90), as.numeric)

data <- data %>% 
  mutate_if(is.character, as.factor)
```


# Participant Portrait

```{r # of participants}
nrow(data)
```


```{r Participant gender}
data %>% 
  group_by(gender) %>% 
  tally() %>%  
  mutate(prop = scales::percent(n/nrow(data)))


data %>%  
  select(gender, gender_other) %>% 
  filter(gender %in% c('Prefer not to answer', 'Prefer to define myself', NA))
# 1 person recognize as agnder, others did not reveal 
```

```{r Participant Age}
range(data$age, na.rm = TRUE)

# 9 missing value  
data %>%  
  ggplot(aes(x = age)) + 
  geom_density() + 
  geom_vline(xintercept = mean(data$age, na.rm = TRUE), color = 'red') + 
  theme_bw()
```

```{r Country of Birth}
# born in Canda
data %>%  
  select(country_of_birth) %>%  
  mutate(Canada = ifelse(country_of_birth == 'Canada', TRUE, FALSE)) %>% 
  group_by(Canada) %>% 
  tally()


data %>%  
  drop_na(country_of_birth) %>% 
  select(country_of_birth) %>% 
  filter(country_of_birth != 'Canada') %>% 
  mutate(continent = case_when(country_of_birth %in% c('China', 'India', 'South Korea', 'Japan', 'Pakistan', 'Philippines') ~ 'Asia', 
                               country_of_birth %in% c('Mauritius', 'Egypt', 'Nigeria') ~ 'Africa',
                               country_of_birth %in% c('United Kingdom of Great Britain and Northern Ireland', 'Czech Republic', 'Italy', 'Germany') ~ 'Europe',
                               country_of_birth %in% c('United States of America', 'Brazil', 'Mexico') ~ 'America')) %>%  
  group_by(continent) %>%  
  tally()
```

```{r Grow-up Province}
data %>% 
  filter(country_of_birth == 'Canada') %>%  
  group_by(grow_up_province) %>% 
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))
```

```{r Field of Study}
data %>%  
  group_by(field_of_study) %>% 
  tally() %>% 
  filter(field_of_study %in% c('Français', 'French', 'Linguistic', 'linguistique et psychologie', 'linguistiques', 'Linguistiques (Arts)'))
# six people are with French or linguitistic background
```
```{r ethnic group}
data %>% 
  group_by(ethnic_group) %>% 
  tally()
```




```{r # of Language Spoken}
# 13 missing data 
data %>% 
  # drop_na(num_of_language) %>% 
  group_by(num_of_language) %>% 
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n))) 
```


```{r Spoken Language}
# 7 missing values 
data %>% 
  drop_na(spoken_language) %>% 
  select(spoken_language, spoken_language_other) %>% 
  mutate(English = ifelse(str_detect(spoken_language, 'English'), TRUE, FALSE)) %>% 
  filter(English != TRUE)
# 2 people don't speak English 
```

```{r Month Tongue}
month_tongue_df <- data %>% 
  select(mother_tongue) %>%
  mutate(English = ifelse(str_detect(mother_tongue, "English") | str_detect(mother_tongue, "Anglais"), TRUE, FALSE),
         Cantonese = ifelse(str_detect(mother_tongue,"Cantonese"), TRUE, FALSE), 
         Hindi = ifelse(str_detect(mother_tongue,"Hindi"), TRUE, FALSE),
         Mandarin = ifelse(str_detect(mother_tongue,"Mandarin"), TRUE, FALSE),      
         Korean = ifelse(str_detect(mother_tongue,"Korean"), TRUE, FALSE),
         Spanish = ifelse(str_detect(mother_tongue,"Spanish"), TRUE, FALSE),  
         German = ifelse(str_detect(mother_tongue,"German"), TRUE, FALSE),  
         Tagalog = ifelse(str_detect(mother_tongue,"Tagalog"), TRUE, FALSE))  

language_count = matrix(NA, nrow = 7, ncol = 2)
language_count[,1] <- names(month_tongue_df)[2:8]
for (i in 1:7){
  language_count[i,2] = as.numeric(table(month_tongue_df[,i+1]))[2]
}


language_count <- as.data.frame(language_count)
colnames(language_count) = c('mother_tongue', 'n')

language_count %>% arrange(desc(n))
```

```{r Year of Studying French}

# 15 missing value
data %>% 
  # drop_na(year_of_french) %>% 
  mutate(year_interval = case_when(year_of_french<=5~'0-5', 
                                   year_of_french>5 & year_of_french <= 10~ '5-10', 
                                   year_of_french>10 & year_of_french<= 15~ '10-15', 
                                   year_of_french > 15~'15+')) %>% 
  group_by(year_interval) %>%
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))

  
  
  
  # drop_
  # tally() %>% 
  # ggplot(aes(x="", y = n, fill=year_group))+
  # geom_bar(width = 1, stat = "identity")+
  # coord_polar("y") + 
  # scale_fill_discrete(name = "Years of Studying French") 
```



```{r Live in French-speaking Area }
live_in_french_env_df <- as.data.frame(t(as.data.frame(data %>% 
  select(french_speaking_place)  %>% 
  group_by(french_speaking_place) %>% 
  tally() %>% 
  mutate(France = ifelse(french_speaking_place == 'France' | french_speaking_place == 'France, Elsewhere? if so please specify:', n, NA),
         Quebec = ifelse(french_speaking_place == 'Quebec' | french_speaking_place == 'Quebec, Elsewhere? if so please specify:', n, NA),
         never = ifelse(str_detect(french_speaking_place, 'I have never stayed in a French-speaking context.'), n, NA),
         other = ifelse(french_speaking_place == 'Elsewhere? if so please specify:', n, NA),
         both = ifelse(french_speaking_place %in% c('Quebec and France', 'Quebec, France'), n, NA),
         unkown = ifelse(is.na(french_speaking_place), n, NA)
         )%>% 
  summarize(France = sum(France, na.rm = TRUE),
            Quebec = sum(Quebec, na.rm = TRUE),
            never = sum(never, na.rm = TRUE),
            other = sum(other, na.rm = TRUE),
            both = sum(both, na.rm = TRUE), 
            unkown = sum(unkown, na.rm = TRUE)))))

live_in_french_env_df <- tibble::rownames_to_column(live_in_french_env_df, "Area")
live_in_french_env_df %>% 
  mutate(prop = scales::percent(V1/sum(V1)))
```
```{r Months of Living in French-speaking areas}
levels(data$month_live_in_fr_env) <- c("1 - 3", "4 - 6", "7 - 12", "< 1", "> 12")
data$month_live_in_fr_env <- factor(data$month_live_in_fr_env, levels = c("< 1", "1 - 3", "4 - 6", "7 - 12", "> 12")) 
data %>% 
  group_by(month_live_in_fr_env) %>% 
  tally() %>%  
  mutate(prop = scales::percent(n/sum(n))) 
```

```{r Preferred Type of French}
data %>% 
  select(pref_french) %>% 
  group_by(pref_french) %>%
  tally() %>% 
  mutate(prop = scales::percent(n/sum(n)))

```


```{r Origin of Teacher}
data %>% 
  group_by(teacher_origin) %>%  
  tally() 

data %>%  
  filter(teacher_origin == 'Other(s), specify:') %>% 
  select(teacher_origin_other)
```


# Part I - Recording Evaluation

1A: Quebec. White.
2A: French. Asian.
3A. Acadian. Asian.
4A. L2 French. White.
5A. African. Asian. 
6A. Quebec. Asian.
7A. African. White. 
8A. Acadian. White. 
9A. French. White. 
10A. L2 French. Asian.




* General Perspective  
1) I think this is beautiful French 


* Understandability
2) Is it rather easy or difficult to understand the speaker?
3) Would the person who did the recording be a good French teacher at UBC?

* Solidarity dimension
5) This person is dynamic.
6) This person is nice.
7) This person is social.

* Status traits
8) This person is professional.
9) This person is leader.
10) This person is educated.


```{r Keep Recording Evluation Columns}
recording_df <- data %>% 
  select(c(1:90)) 
```

```{r Beautiful French}
beautiful_fr <- 
  recording_df %>% 
  select(seq(1, 82, 9)) %>% 
  transmute(Quebec = `1A_1` + `6A_1`, 
         French = `2A_1` + `9A_1`, 
         Acadian = `3A_1` + `8A_1`, 
         `L2 French` = `4A_1` + `10A_1`, 
         African = `5A_1` + `7A_1`
        ) %>% 
  mutate(id = 1:nrow(recording_df))


# short to long 
beautiful_fr_long <- beautiful_fr %>%  
  gather(key = "Accent", value = "value", Quebec, French, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, Accent)


beautiful_fr_long %>%  
  group_by(Accent) %>% 
  get_summary_stats(value, type = 'mean_sd')
# Note: only 83 people complete the survey



# repeated measure ANOVA
anova_beau_fr <-
  get_anova_table(anova_test(
    data = beautiful_fr_long,
    dv = value,
    wid = id,
    within = Accent
  ))

# paired t test
beautiful_fr_pw <- beautiful_fr_long %>%
  pairwise_t_test(value ~ Accent,
                  paired = TRUE,
                  p.adjust.method = 'bonferroni') %>% 
  mutate(y.position = c(seq(12, 16.5, by = 0.5)))


p_beautiful <- ggboxplot(beautiful_fr_long, x = "Accent", y = "value", fill = "Accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(beautiful_fr_pw, vjust = 0.1, bracket.nudge.y = 0.6, size = 2,hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(anova_beau_fr, detailed = TRUE),
    caption = get_pwc_label(beautiful_fr_pw),
    title = 'Beautiful French',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))

```


1A: Quebec. White.
2A: French. Asian.
3A. Acadian. Asian.
4A. L2 French. White.
5A. African. Asian. 
6A. Quebec. Asian.
7A. African. White. 
8A. Acadian. White. 
9A. French. White. 
10A. L2 French. Asian.


```{r Solidarity dimension - create df}
# 5) This person is dynamic.
# 6) This person is nice.
# 7) This person is social.

sd_df <- recording_df %>% 
  select(contains(c('_5', '_6', '_7'))) %>%
  transmute(que_dynamic = `1A_5` + `6A_5`, 
         fr_dynamic = `2A_5` + `9A_5`, 
         aca_dynamic = `3A_5` + `8A_5`, 
         l2_dynamic = `4A_5` + `10A_5`, 
         afr_dynamic = `5A_5` + `7A_5`, 
         que_nice = `1A_6` + `6A_6`, 
         fr_nice = `2A_6` + `9A_6`, 
         aca_nice = `3A_6` + `8A_6`, 
         l2_nice = `4A_6` + `10A_6`, 
         afr_nice = `5A_6` + `7A_6`, 
         que_social = `1A_7` + `6A_7`, 
         fr_social = `2A_7` + `9A_7`, 
         aca_social = `3A_7` + `8A_7`, 
         l2_social = `4A_7` + `10A_7`, 
         afr_social = `5A_7` + `7A_7`, 
          )

# dynamic 
dynamic_df_long <- sd_df %>% 
  select(c(1:5)) %>%  
  rename(Quebec = que_dynamic,
         European = fr_dynamic,
         Acadian = aca_dynamic,
         `L2 French` = l2_dynamic,
         African = afr_dynamic
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)


# nice 
nice_df_long <- sd_df %>% 
  select(c(6:10)) %>%  
  rename(Quebec = que_nice,
         European = fr_nice,
         Acadian = aca_nice, 
         `L2 French` = l2_nice,
         African = afr_nice
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)

# social 

social_df_long <- sd_df %>% 
  select(c(11:15)) %>%  
  rename(Quebec = que_social,
         European = fr_social,
         Acadian = aca_social,
         `L2 French` = l2_social,
         African = afr_social
         ) %>% 
  mutate(id = 1: nrow(sd_df)) %>%  
  gather(key = "accent", value = "value", Quebec, European, Acadian, `L2 French`, African) %>%
  convert_as_factor(id, accent)
```

```{r SD - Dynamic}
dynamic_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_dynamic <- get_anova_table(anova_test(data = dynamic_df_long, dv = value, wid = id, within = accent))
  
dynamic_pw <- dynamic_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1))) 


p_dynamic <- ggboxplot(dynamic_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(dynamic_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_dynamic, detailed = TRUE),
    caption = get_pwc_label(dynamic_pw),
    y = 'Score',
    title = 'Dynamic'
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))
```
```{r SD - Nice}
nice_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  

res_nice <- get_anova_table(anova_test(data = nice_df_long, dv = value, wid = id, within = accent))
  
nice_pw <- nice_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1)))
p_nice <- ggboxplot(nice_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(nice_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) + 
  labs(
    subtitle = get_test_label(res_nice, detailed = TRUE),
    caption = get_pwc_label(nice_pw),
    title = 'Nice',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold'))

```

```{r SD - Social}
social_df_long %>% 
  group_by(accent) %>%
  get_summary_stats(value, type = "mean_sd")  
res_social <- get_anova_table(anova_test(data = social_df_long, dv = value, wid = id, within = accent))
  
social_pw <- social_df_long %>%
  pairwise_t_test(
    value ~ accent, paired = TRUE,
    p.adjust.method = "bonferroni"
    ) %>% 
  mutate(y.position = c(seq(12,21, by = 1)))
(p_social <- ggboxplot(social_df_long, x = 'accent', y = "value", fill = "accent") + 
  rremove('legend') + 
  rremove('xlab') + 
  stat_pvalue_manual(social_pw, vjust = 0.5, bracket.nudge.y = 0.5, size = 3, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res_social, detailed = TRUE),
    caption = get_pwc_label(social_pw),
    title = 'Social',
    y = 'Score',
  ) + 
  theme(plot.caption = element_text(size=12),
        plot.subtitle=element_text(size=13, vjust= 1, face="italic", color="black"),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size=15, face = 'bold')))
```

```{r SD Results, fig.height= 8, fig.width= 8}
cowplot::plot_grid(p_beautiful,p_dynamic, p_nice, p_social)

```


```{r}
test_df_2 <- beautiful_fr %>% 
  transmute(Quebec = `1A_1` + `6A_1`, 
         French = `2A_1` + `9A_1`, 
         Acadian = `3A_1` + `8A_1`, 
         L2_French = `4A_1` + `10A_1`, 
         African = `5A_1` + `7A_1`
        )
```




```{r }
names(recording_df)
```

#
















